We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2017-11-26 paper/paper.tex
new results from illustrisdm

\begin{abstract}
Here we quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We build explicit probability distributions of finding an aspherical
satellite distribution. 
This allows us to estimate the atipicallity of the satellite
distributions in the LG even when the underlying simulations
do not have enough systems that resemble the LG features.
This is also standarized method to compare different simulations and
is also convenient for reproducibility and comparison 
We demonstrate the method using the Illustris-1 and ELVIS simulations.
We find significant differences between the two simulations. 
In the case of Illustris-1 $0.05\%$ to $0.25\%$ of the pairs with isolation
and dynamical characteristics similar to the LG are expected to have
satellite distributions with the same degree of deviation from
sphericity.
In the ELVIS simulations this fractions drops dramatically to
$0.001\%$ of the pairs.
This makes the LG a $3\sigma$ outlier in LCDM simulations, with all
the weight lying on the extremly planar MW satellite distribution.

@@

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{expected_numbers.pdf}
\includegraphics[width=0.47\textwidth]{expected_numbers_flip.pdf}
\caption{Probability distribution for the expected number of pairs
  showing the same degree of atipicality as the Local Group if drawn
  from a sample of 2000 isolated pairs. 
  The distributions correspond to results derived from Illustris-1
  and ELVIS data.
  On average between $0.05\%$ and $0.25\%$ of the isolated pairs should present
  satellite distributions as atypical as the Local Group.
\label{fig:expected_number}}
\end{figure}

>>>>

\begin{abstract}
Here we quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We build explicit probability distributions of finding an aspherical
satellite distribution. 
This allows us to estimate the atipicallity of the satellite
distributions in the LG even when the underlying simulations
do not have enough systems that resemble the LG features.
This is also standarized method to compare different simulations and
is also convenient for reproducibility and comparison 
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-DarkMatter and ELVIS simulations.
Although there are significant differences between the three
simulations, we find that at most of $0.20\%$ the pairs are expected
to have satellite distributions with the same degree of deviation from
sphericity as the Local Group.
The three simulations show that the satellite
distribution in M31 is common while the very planar MW satellite
distribution is hard to find; we expect that between $19\%$ to $24\%$
of the pairs have halo similar to M31, while only $0.80\%$ are similar
to the MW.

@@

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{expected_numbers_n_M31.pdf}
\includegraphics[width=0.47\textwidth]{expected_numbers_n_MW.pdf}
\caption{Probability distribution for the expected number of pairs
  showing the same degree of atipicality as the Local Group if drawn
  from a sample of 2000 isolated pairs. 
  The distributions correspond to results derived from Illustris-1
  and ELVIS data.
  On average between $0.05\%$ and $0.25\%$ of the isolated pairs should present
  satellite distributions as atypical as the Local Group.
\label{fig:expected_number}}
\end{figure}
